---
title: Bee foraging distances dependent on body mass and level of eusociality - a
  meta-analysis
author: Kendall, L.K.^1^, Mola, J. ^2^, Bartomeus, I.^3^,Cariveau, D. ^4^, Portman, Z.^4^ (TBC)
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
  word_document: default
html_document: default
editor_options: null
chunk_output_type: console
---

*Affiliations:*

^1^ Centre for Environmental and Climate Science, Lunds Universitet, Lund, Sweden 
^2^ U.S.D.A. Geological Survey? 
^3^ Estación Biológica de Doñana (EBD-CSIC), Avda. Américo Vespucio 26, Isla de la Cartuja, E-41092 Sevilla, Spain   
^4^ University of Minnesota

*Correspondence: liam.k.kendall@gmail.com   

```{r setup dataframes, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#libraries
library(brms)
library(plyr)
library(tidyverse)
library(glmmTMB)
library(DHARMa)
library(ggplot2)
library(pollimetry)
library(phytools)
library(ape)
library(performance)

##dataframes
distances <- read.csv("data/species_distances_final.csv",
                      sep=";",dec=",",
                      stringsAsFactors = FALSE)
colnames(distances)[1] = "Authors"

distances$spp<-distances$Genus_species
distances$obs<-1:nrow(distances)

#merge with traits
traits <- read.csv("data/species_traits.csv",
                   sep=";",dec=",",
                   stringsAsFactors = FALSE)

data("pollimetry_dataset")
traits[traits$Genus_species%in%"Megachile_femorata","ITD"] = mean(pollimetry_dataset[pollimetry_dataset$Genus%in%"Megachile" & pollimetry_dataset$Sex%in%"Female","IT"])

#M. femorata = itd = average of other Megachile from pollimetry dataset

#log distance and ITD
distances$log.dist<-log(distances$Distance)
traits$log.it<-log(traits$ITD)


distances$Genus<-word(distances$Genus_species,1,sep="_")

distances$limits <- ifelse(distances$Metric2%in%c("feeder","homing"),"physio","funct")
#join
forage.traits <- distances%>%
  left_join(traits[,-1],by=c("Genus_species"))%>%
  filter(!is.na(ITD)==T)%>%#minus 3 
  filter(!Sex_Imp%in%c("Male"))%>%#minus 3 
  filter(!Notes1%in%c("dispersal distance"))%>%
  filter(!Metric2%in%c(NA))%>%
  filter(!distance_type2%in%c("Observed"))%>%
  mutate(Sex_Imp=revalue(Sex_Imp,c("Female_Queen" = "Female")))#%>%
  #filter(!distance_type%in% 
  #c("Minimum"))%>%
  #filter(!is.na(Metric2)==T)

#observed distances
forage.traits.obs <- distances%>%
  left_join(traits[,-1],by=c("Genus_species"))%>%
  filter(!is.na(ITD)==T)%>%#minus 3 
  filter(!Sex_Imp%in%c("Male"))%>%#minus 3 
  filter(!Notes1%in%c("dispersal distance"))%>%
  filter(!Metric2%in%c(NA))%>%
  filter(distance_type2%in%c("Observed"))

#Removed males, queens
#Minimum is pollen mapping study with "minimum distance to a given pollen source" = incompatible

#Aggregate all mean / median / mode / typical estimates
#forage.traits$distance_type2 <- 
#  revalue(forage.traits$distance_type,
#          c("max_50" = "Central", #homing 50% 
##            "Mode" = "Central",
##            "Mean" = "Central",
#            "Median" = "Central",
#            "Typical" = "Central",
##            "max_90" = "Max", #90% 
#            "max_95" = "Max")) #95% homing 

#max dataframe 158
forage.traits.max <- forage.traits%>%
  filter(distance_type2%in%"Max")

#mean dataframe 111
forage.traits.mean <- forage.traits%>%
  filter(distance_type2%in%"Typical")

#max and mean solitary
forage.traits.max.sol<-forage.traits.max%>%
  filter(social_tree%in%"Solitary")%>%
  filter(!Lecty%in%c("unclear",NA))

forage.traits.mean.sol<-forage.traits.mean%>%
  filter(social_tree%in%"Solitary")%>%
  filter(!Lecty%in%c("unclear",NA))

##phylo setup (same as pollimetry)

#tree
bee.trees=read.tree(file="data/hedtke_genera_tree.txt")

##Use tree 1 (376 genera) #Genera-level phylogney
bee.mcmc=bee.trees[[1]]%>%
  root(outgroup="Tachysphex")%>%
  as.phylo

bee.cal<-makeChronosCalib(bee.mcmc)
bee.mcmc=chronos(bee.mcmc,calibration = bee.cal)

bee.max.tree=drop.tip(bee.mcmc, setdiff(bee.mcmc$tip.label,
                                        unique(forage.traits.max$Genus)))

bee.mean.tree=drop.tip(bee.mcmc, setdiff(bee.mcmc$tip.label,
                                         unique(forage.traits.mean$Genus)))

bee.max.tree$node.label=NULL
bee.mean.tree$node.label=NULL

#load models
setwd("./model_outputs/")
file_names=as.list(dir(pattern=".rdata"))
for(i in 1:length(file_names)) load(file_names[[i]])

#list model objects
mod.list <- list(max.IT.brm,
max.euc.phylo.full.brm,
max.euc.phylo.reduced.brm,
max.func.phylo.brm,
max.func.euc.full.phylo.brm,
max.func.euc.red.phylo.brm,
max.lecty.phylo.brm.1,
max.lecty.phylo.brm.2,
mean.IT.brm,
mean.euc.phylo.full.brm,
mean.euc.phylo.reduced.brm,
mean.func.phylo.brm,
mean.func.euc.full.phylo.brm,
mean.func.euc.red.phylo.brm,
mean.metric.phylo.brm,
mean.lecty.phylo.brm.1,
mean.lecty.phylo.brm.2)

names(mod.list) <- c("maxIT","maxEUF","maxEUR","maxFUNCsimple","maxFUNCeusFULL","maxFUNCeusRED","maxLEF","maxLER",
                     "meanIT","meanEUF","meanEUR","meanFUNCsimple","meanFUNCeusFULL","meanFUNCeusRED",
                         "meanMET","meanLEF","meanLER")

model.r2s <- do.call("rbind", lapply(mod.list,function(x) r2(x)))

```

##THOUGHTS - while writing and reading

* Is distance the right word? I cant decide between distance or range, range seems more synonymous with resource use and therefore maybe better i.e. they will forage within this range, rather than less than this distance

* I think the physiological ~ functional split in estimates (feeder and homing vs. all others) is very interesting. I am open to ideas regarding the names - maybe behavioural is better than functional? 

* I didn't find an effect of polylecty for either maximum or typical distances - i thought there was one but there were a few estimates which were incorrectly coded- and after fixing these, we lost the effect.
Two options: we focus just on eusociality, as the result is strong, or add the lecty does not have an effect in solitary species?
Lecty results are not here at present, but model objects are - have lecty in the name i.e. (max.lecty.phylo.brm.1 etc)

* Not enough coverage to use Family as a predictor (see table(forage.traits$Family,forage.traits$distance_type2))

* Henrik had an interesting idea - use colony size instead of categorical eusociality


##IDEAS
Include measure of variance in intraspecific ITD into models as measurement error

# Abstract

# Introduction

Paragraph One: Ecological significance of foraging in terms of ecosystem function.
Effects on foraging - landscape context etc
-links to body size

Types of measurements
Structure that feeder and homing as physiological measures - others as functional

Why should this be revisited? 
-Many more studies
-new analytical methods
-need to incorporate things like phylogeny etc

##Research questions 

1) How accurate is body size in predicting foraging distance in bees?

2) Do bees differ in terms of their physiological and functional constraints in foraging distance/range?

2) i) Does level of eusociality nfluence foraging distance?
  ii) Do the functional and physiological constraints on foraging distance differ between eusocial and solitary bee species?

3) Does floral specialisation structure species-level foraging distance in solitary bee species?

# Materials and Methods

To identify relevant publications for our meta-analysis, we compiled foraging distance estimates from a previous meta-analysis by Greenleaf et al. 2007 and a litearture review by Zurbuchen et al. 2010. We then undertook a systematic Scopus database search with the following search terms: TITLE-ABS-KEY ( ( bee  OR  apoid* )  AND  ( forag*  AND  distanc* )  OR  "flight range"). As of 12 March 2021, this returned 643 articles. We then subsetted this database to articles published post-2006 (to coincide with Greenleaf et al. 2007) and screened all articles for foraging distance estimates (n = 180 articles). For each estimate, we delimited estimate by type of distances measured (maximum, typical or observed) and methodology used to measure foraging distance (see Table 1 for definitions), following Greenleaf et al. 2007. We focussed on female bees, excluding all species that focussed on males or specifically dispersal patterns.

```{r definition table, echo=FALSE}
text_tbl <- data.frame(
  Group = c("Type of distance","","",
            "Method","","","","","",""),
  Type = c("Typical","Maximum","Observed","Functional","",
            "","","","Physiological",""),
  Type2 = c("","","","Associative","Mark-recapture",
            "Molecular","Tracking","Waggle dance","Feeder","Homing"),
  Definition = c(
    "Typical foraging distance was defined as the mean (or median or mode) foraging distance by a species within a study.",
    "Maximum foraging distance was defined asthe maximum distance observed within a experiment (if multiple trials were conducted) or a study",
    "Observed foraging distances were those estaimtes that could not be definied as either typical or maximum distances i.e....","Associative studies were those that measured foraging distance by associating known nests / hives and floral resources..",
    "Mark-recapture...",
        "Molecular...",
        "Tracking...",
    "Dancing...",
    "Feeder studies involved...",
    "Homing studies..."))
library(kableExtra)
kbl(text_tbl) %>%
  kable_paper(full_width = F) %>%
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(2, bold = T, border_right = T) %>%
  column_spec(3, bold = T, border_right = T) %>%
  column_spec(4, width = "30em")
```
**Table 1.** Types of bee foraging estimates and methodologies..

##Species taxonomy and functional traits

We checked the validity of species' names using X, Y, and Z. We then collected the following traits for each species: female body size (measured as the intertegular distance, Cane et al. 1987), level of eusociality and floral specialisation. 
We obtained body size from literature sources (i.e. pollimetry, brazilian trait dataset, traitbase) and by directly contacting researchers. 
Where multiple ITD measurements for a given species were found, we took the mean value. 
For eusocial species with castes, we obtained ITD measurements for the caste used within the study (i.e. workers or queens).

We delimited species into three levels of sociality: *obligate solitary*, *primitively eusocial* or *highly eusocial*. Obligate solitary species were those bees species that provision their own nest cells. This group includes species that live communally or in aggregations. 
Primitively eusocial species were those species that have non-permanent colonies and that are founded by a single female. This group included bumblebees *Bombus* species, as well as socially flexible *Lasioglossum* species, and those with overlapping generations e.g. *Xylocopa* species.
Highly eusocial species were those species that had permanent colonies, which reproduce by budding. This group included all honeybees *Apis* species, and the stingless bees (Meliponini).

Lastly, we grouped species,based on their level of floral specialisation, as either *oligolectic* or *polylectic*.
Oligolectic speies were defined as those species who consistently collect pollen from a single plant species or group of similar or related plant species, even the presene of other pollen sources. 
Polylectic species were defined as those that use multiple, unrelated species for pollen.

Nesting? I have not done any analysis on this (yet). 

##Statistical analysis

All statistical analyses were undertaken in *R* (citation).

We used Bayesian generalized linear mixed effect models to analyze the effect of body size, functional traits and study design on bee foraging distances.
We modeled typical and maximum foraging distance separately.
Observed foraging distances were not analysed. Furhter, we excluded all estimates within unknown metholodogy.
As is typical in allometric studies, we log-transformed foraging distance prior to analyses. However, as intertegular distance is logarithmically related to body size (Cane et al. 1987; Kendall et al. 2019), we compared the predictive accuracy (R^2) of models with and without log-transformed ITD. We present models with the raw ITD, as these were significantly more predictive than ln(ITD). 

For these purposes we specified the following models:
1) Foraging distance as a function of ITD, 2) foraging distance as a function of ITD * eusociality (categorical, three levels), using the complete dataset, 3) foraging distance as a function of ITD * Lecty in solitary bees species, as oligolecty is only found within solitary species and 4) foraging distance as a function of ITD * Metric type (physiological or functional) * eusociality. Floral specialisation was not tested in conjunction with metric type due to small sample sizes.

For models 1, 2 and 3), the random effect structure consisted of intercepts for publication and bee species, a phylogenetic effect, which consisted of a phylogenetic relatedness correlation matrix between bee genera, and a random intercept and slope for metric. Model (4) had the same random structure but without the metric effect as this was modelled as a fixed effect. For the phylogenetic effect, we constructed a bee phylogeny using an available bee genera tree (Hedtke, Patiny, & Danforth, 2013). We removed non-represented genera and then fitted a chronogram from this phylogeny by penalized likelihood, and a correlated rate model (Paradis et al. 2004). We then converted this to a phylogenetic correlation matrix using *ape* package.

We assessed the significance of the main- and interactive effects in each models by assessing whether 95% credible intervals bracketed zero. We then removed unsignificant interactions.  In model (i), the interaction between ITD and eusociality was non-significant for both typical and maximum foraging distance, so we fitted and present the reduced models. 

We parameterised each model with a *Gaussian* distribution, which was run for 2000 iterations, and a burn-in of 1000 iterations. We set weakly informative priors and manipulated DELTA and maximum tree depth to avoid divergent transitions. We visualised posterior predictive checks using bayesplot (v1.6.0, Gabry & Mahr 2017) and assessed chain convergence for each sub-model separately. We considered effects were considered significant when 95% credible intervals did not bracket zero. 

# Results
```{r All max models, eval = FALSE, echo=FALSE,include=FALSE}
#DOES NOT RUN 
#here to see model formulations

#run models
max.priors <- prior(normal(5,2),class="Intercept")+
  prior(normal(-1,1),class="b")+
  prior(normal(0,1),class="b",coef="ITD")+
  prior(normal(0,1.25),class="sd")+
  prior(normal(0,1),class="sigma")

max.IT.brm <- brm(log.dist~ITD+
                             (1|Authors)+
                             (ITD|Metric2)+
                             (1|spp)+
                             (1|gr(Genus, cov = Ag)),
                           data2 = list(Ag = Ag),
                           cores=4,
                           iter = 2000,
                           prior = max.priors,
                           family="gaussian",
                           control = list(adapt_delta=0.999,
                                          max_treedepth=15),
                           data=forage.traits.max)

max.euc.phylo.full.brm <- brm(log.dist~ITD*social_tree+
                             (1|Authors)+
                             (ITD|Metric2)+
                             (1|spp)+
                             (1|gr(Genus, cov = Ag)),
                           data2 = list(Ag = Ag),
                           cores=4,
                           iter = 2000,
                           prior = max.priors,
                           family="gaussian",
                           control = list(adapt_delta=0.999,
                                          max_treedepth=15),
                           data=forage.traits.max)

max.euc.phylo.reduced.brm <- brm(log.dist~ITD+social_tree+
                               (1|Authors)+
                               (ITD|Metric2)+
                            (1|spp)+
                            (1|gr(Genus, cov = Ag)),
                          iter = 2000,
                          prior = max.priors,
                          data2 = list(Ag = Ag),
                          cores=4,
                          family="gaussian",
                          control = list(adapt_delta=0.999,
                                         max_treedepth=15),
                          data=forage.traits.max)

#func vs physio models

max.func.priors <- prior(normal(5,2),class="Intercept")+
  prior(normal(-1,1),class="b")+
  prior(normal(0,1),class="b",coef="ITD")+
  prior(normal(0,1),class="b",coef="ITD:limitsphysio")+
  prior(normal(0,1.25),class="sd")+
  prior(normal(0,1),class="sigma")

#physio vs functional model
max.func.phylo.brm<- brm(log.dist~ITD*limits+
                             (1|Authors)+
                             (ITD|Metric2)+
                             (1|spp)+
                             (1|gr(Genus, cov = Ag)),
                           data2 = list(Ag = Ag),
                           cores=4,
                           iter = 2000,
                           prior = max.func.priors,
                           family="gaussian",
                           control = list(adapt_delta=0.999,
                                          max_treedepth=15),
                           data=forage.traits.max)

max.func.euc.full.phylo.brm<- brm(log.dist~ITD*limits*social_tree+
                           (1|Authors)+
                           (ITD|Metric2)+
                           (1|spp)+
                           (1|gr(Genus, cov = Ag)),
                         data2 = list(Ag = Ag),
                         cores=4,
                         iter = 2000,
                         prior = max.func.priors,
                         family="gaussian",
                         control = list(adapt_delta=0.999,
                                        max_treedepth=15),
                         data=forage.traits.max)

max.func.euc.red.phylo.brm<- brm(log.dist~ITD*limits+social_tree+
                             (1|Authors)+
                             (ITD|Metric2)+
                             (1|spp)+
                             (1|gr(Genus, cov = Ag)),
                           data2 = list(Ag = Ag),
                           cores=4,
                           iter = 2000,
                           prior = max.func.priors,
                           family="gaussian",
                           control = list(adapt_delta=0.999,
                                          max_treedepth=15),
                           data=forage.traits.max)


#metric model
max.metric.phylo.brm<- brm(log.dist~ITD*Metric2+
                            (1|Authors)+
                            (1|spp)+
                            (1|gr(Genus, cov = Ag)),
                          data2 = list(Ag = Ag),
                          cores=4,
                          iter = 2000,
                          prior = max.priors,
                          family="gaussian",
                          control = list(adapt_delta=0.999,
                                         max_treedepth=15),
                          data=forage.traits.max)

#lecty models
bee.max.sol.tree<-drop.tip(bee.max.tree,setdiff(bee.max.tree$tip.label,
                                                forage.traits.max.sol$Genus))
bee.max.sol.tree$node.label=NULL

Mlg <- ape::vcv.phylo(bee.max.sol.tree)

forage.traits.max.sol$Lecty <- relevel(factor(forage.traits.max.sol$Lecty),ref="Polylectic")


max.lecty.priors <- prior(normal(5,2),class="Intercept")+
  prior(normal(0,2),class="b")+
  prior(normal(0,1),class="b",coef="ITD")+
  prior(normal(0,1),class="sd")+
  prior(normal(0,1),class="sigma")

max.lecty.phylo.brm.1<- brm(log.dist~ITD*Lecty+
                              (ITD|Metric2)+
                              (1|Authors)+
                              (1|spp)+
                              (1|gr(Genus, cov = Mlg)),
                            prior = max.lecty.priors,
                            iter = 4000,
                            data2 = list(Mlg = Mlg),
                            cores=4,
                            family="gaussian",
                            control = list(adapt_delta=0.999,
                                           max_treedepth=15),
                            data=forage.traits.max.sol)

max.lecty.phylo.brm.2<- brm(log.dist~ITD+Lecty+
                              (ITD|Metric2)+
                              (1|Authors)+
                              (1|spp)+
                              (1|gr(Genus, cov = Mlg)),
                            prior = max.lecty.priors,
                            iter = 4000,
                            data2 = list(Mlg = Mlg),
                            cores=4,
                            family="gaussian",
                            control = list(adapt_delta=0.999,
                                           max_treedepth=15),
                            data=forage.traits.max.sol)

#save all mdoel objects
save(max.IT.brm,file="model_outputs/max.IT.brm.rdata")
save(max.euc.phylo.full.brm,file="model_outputs/max.euc.phylo.full.brm.rdata")
save(max.euc.phylo.reduced.brm,file="model_outputs/max.euc.phylo.reduced.brm.rdata")
save(max.metric.phylo.brm,file="model_outputs/max.metric.phylo.brm.rdata")
save(max.lecty.phylo.brm.1,file="model_outputs/max.lecty.phylo.brm.1.rdata")
save(max.lecty.phylo.brm.2,file="model_outputs/max.lecty.phylo.brm.2.rdata")
save(max.func.phylo.brm,file="model_outputs/max.func.phylo.brm.rdata")
save(max.func.euc.full.phylo.brm,file="model_outputs/max.func.euc.full.phylo.brm.rdata")
save(max.func.euc.red.phylo.brm,file="model_outputs/max.func.euc.red.phylo.brm.rdata")


```

```{r All mean typical models, eval = FALSE, echo=FALSE,include=FALSE}

#####
mean.priors <- prior(normal(5,2),class="Intercept")+
  prior(normal(-1,1),class="b")+
  prior(normal(0,1),class="b",coef="ITD")+
  prior(normal(0,1),class="sd")+
  prior(normal(0,1),class="sigma")


mean.IT.brm<- brm(log.dist~ITD+
                            (1|Authors)+
                            (ITD|Metric2)+
                            (1|spp)+
                            (1|gr(Genus, cov = Bg)),
                          iter = 2000,
                          prior = mean.priors,
                          data2 = list(Bg = Bg),
                          cores=4,
                          family="gaussian",
                          control = list(adapt_delta=0.999,
                                         max_treedepth=15),
                          data=forage.traits.mean)

mean.euc.phylo.full.brm <- brm(log.dist~ITD*social_tree+
                            (1|Authors)+
                            (ITD|Metric2)+
                            (1|spp)+
                            (1|gr(Genus, cov = Bg)),
                          iter = 2000,
                          prior = mean.priors,
                          data2 = list(Bg = Bg),
                          cores=4,
                          family="gaussian",
                          control = list(adapt_delta=0.999,
                                         max_treedepth=15),
                          data=forage.traits.mean)

mean.euc.phylo.reduced.brm <- brm(log.dist~ITD+social_tree+
                            (1|Authors)+
                            (ITD|Metric2)+
                            (1|spp)+
                            (1|gr(Genus, cov = Bg)),
                          iter = 2000,
                          prior = mean.priors,
                          data2 = list(Bg = Bg),
                          cores=4,
                          family="gaussian",
                          control = list(adapt_delta=0.999,
                                         max_treedepth=15),
                          data=forage.traits.mean)

##func vs physio models

#func model
mean.func.priors <- prior(normal(5,2),class="Intercept")+
  prior(normal(-1,1),class="b")+
  prior(normal(0,1),class="b",coef="ITD:limitsphysio")+
  prior(normal(0,1),class="b",coef="ITD")+
  prior(normal(0,1),class="sd")+
  prior(normal(0,1),class="sigma")
#physio vs functional model
mean.func.phylo.brm<- brm(log.dist~ITD*limits+
                           (1|Authors)+
                           (ITD|Metric2)+
                           (1|spp)+
                           (1|gr(Genus, cov = Ag)),
                         data2 = list(Ag = Ag),
                         cores=4,
                         iter = 2000,
                         prior = mean.func.priors,
                         family="gaussian",
                         control = list(adapt_delta=0.999,
                                        max_treedepth=15),
                         data=forage.traits.mean)

mean.func.euc.full.phylo.brm<- brm(log.dist~ITD*limits*social_tree+
                                    (1|Authors)+
                                    (ITD|Metric2)+
                                    (1|spp)+
                                    (1|gr(Genus, cov = Ag)),
                                  data2 = list(Ag = Ag),
                                  cores=4,
                                  iter = 2000,
                                  prior = mean.func.priors,
                                  family="gaussian",
                                  control = list(adapt_delta=0.999,
                                                 max_treedepth=15),
                                  data=forage.traits.mean)

mean.func.euc.red.phylo.brm<- brm(log.dist~ITD*limits+social_tree+
                                   (1|Authors)+
                                   (ITD|Metric2)+
                                   (1|spp)+
                                   (1|gr(Genus, cov = Ag)),
                                 data2 = list(Ag = Ag),
                                 cores=4,
                                 iter = 2000,
                                 prior = mean.func.priors,
                                 family="gaussian",
                                 control = list(adapt_delta=0.999,
                                                max_treedepth=15),
                                 data=forage.traits.mean)


#metric model
mean.metric.phylo.brm<- brm(log.dist~ITD*Metric2+
                             (1|Authors)+
                             (1|spp)+
                             (1|gr(Genus, cov = Bg)),
                           data2 = list(Bg = Bg),
                           cores=4,
                           iter = 2000,
                           prior = mean.priors,
                           family="gaussian",
                           control = list(adapt_delta=0.999,
                                          max_treedepth=15),
                           data=forage.traits.mean)

mean.metric.phylo.brm.2<- brm(log.dist~ITD+Metric2+
                               (1|Authors)+
                               (1|spp)+
                               (1|gr(Genus, cov = Bg)),
                             data2 = list(Bg = Bg),
                             cores=4,
                             iter = 2000,
                             prior = mean.priors,
                             family="gaussian",
                             control = list(adapt_delta=0.999,
                                            max_treedepth=15),
                             data=forage.traits.mean)

#mean

bee.mean.sol.tree<-drop.tip(bee.mean.tree,setdiff(bee.mean.tree$tip.label,
                                                forage.traits.mean.sol$Genus))
bee.mean.sol.tree$node.label=NULL
plot(bee.mean.sol.tree)

Melg <- ape::vcv.phylo(bee.mean.sol.tree)

forage.traits.mean.sol$Lecty <- relevel(factor(forage.traits.mean.sol$Lecty),ref="Polylectic")

mean.lecty.priors <- prior(normal(0,5),class="Intercept")+
  prior(normal(0,2),class="b")+
  #prior(normal(0,1),class="b",coef="ITD")+
  prior(normal(0,1),class="sd")+
  prior(normal(0,1),class="sigma")

mean.lecty.phylo.brm.1<- brm(log.dist~ITD*Lecty+
                              (ITD|Metric2)+
                              (1|Authors)+
                              (1|spp)+
                              (1|gr(Genus, cov = Melg)),
                            prior = mean.lecty.priors,
                            iter = 4000,
                            data2 = list(Melg = Melg),
                            cores=4,
                            family="gaussian",
                            control = list(adapt_delta=0.999,
                                           max_treedepth=15),
                            data=forage.traits.mean.sol)

mean.lecty.phylo.brm.2<- brm(log.dist~ITD+Lecty+
                              (ITD|Metric2)+
                              (1|Authors)+
                              (1|spp)+
                              (1|gr(Genus, cov = Melg)),
                            prior = mean.lecty.priors,
                            iter = 4000,
                            data2 = list(Melg = Melg),
                            cores=4,
                            family="gaussian",
                            control = list(adapt_delta=0.999,
                                           max_treedepth=15),
                            data=forage.traits.mean.sol)

##save all model objects
save(mean.IT.brm,file="model_outputs/mean.IT.brm.rdata")
save(mean.euc.phylo.full.brm,file="model_outputs/mean.euc.phylo.full.brm.rdata")
save(mean.euc.phylo.reduced.brm,file="model_outputs/mean.euc.phylo.reduced.brm.rdata")
save(mean.metric.phylo.brm,file="model_outputs/mean.metric.phylo.brm.rdata")
save(mean.lecty.phylo.brm.1,file="model_outputs/mean.lecty.phylo.brm.1.rdata")
save(mean.lecty.phylo.brm.2,file="model_outputs/mean.lecty.phylo.brm.2.rdata")
save(mean.func.phylo.brm,file="model_outputs/mean.func.phylo.brm.rdata")
save(mean.func.euc.full.phylo.brm,file="model_outputs/mean.func.euc.full.phylo.brm.rdata")
save(mean.func.euc.red.phylo.brm,file="model_outputs/mean.func.euc.red.phylo.brm.rdata")

```

##Summary

In total, we included `r (nrow(forage.traits.max)+nrow(forage.traits.mean))` estimates of foraging distance (typical, n = `r nrow(forage.traits.mean)`) & maximum, n = `r nrow(forage.traits.max)`) in our analyses. These included estimates from `r length(levels(factor(forage.traits$Genus_species)))` species, `r length(levels(factor(forage.traits$Genus)))` genera and `r length(levels(factor(forage.traits$Family)))` families. Honeybees *Apis* spp. (`r nrow(forage.traits[forage.traits$Genus%in%c("Apis"),])` estimates) and bumblebees *Bombus* spp. (`r nrow(forage.traits[forage.traits$Genus%in%c("Bombus"),])` estimates) were the most commonly studied taxa. Maximum foraging distance estimates ranged from `r min(forage.traits.max$Distance)` metres (*`r gsub("_"," ",forage.traits.max[forage.traits.max$Distance%in%min(forage.traits.max$Distance),"Genus_species"])`*) to `r max(forage.traits.max$Distance)/1000` kilometres (*`r gsub("_"," ",forage.traits.max[forage.traits.max$Distance%in%max(forage.traits.max$Distance),"Genus_species"])`*). Typical foraging distance estimates ranged from `r min(forage.traits.mean$Distance)` metres (*`r gsub("_"," ",forage.traits.mean[forage.traits.mean$Distance%in%min(forage.traits.mean$Distance),"Genus_species"])`*) to `r max(forage.traits.mean$Distance)/1000` kilometres (*`r gsub("_"," ",forage.traits.mean[forage.traits.mean$Distance%in%max(forage.traits.mean$Distance),"Genus_species"])`*).

ADD TABLE OF ESTIMATES -species summary

In addition, we compiled `r nrow(distances%>%filter(distance_type2%in%"Observed"))` 'observed' foraging distance estimates that were not analysed. 

```{r figure 1, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
library(patchwork)
library(ggplot2)
#library(Manu) NZ native bird palettes

col.palette <- c("#325756","#7d9fc2","#C582B2","#51806a","#4d5f8e","#A092B7")

#Fitted points from models 
#overall
overall.pred <- rbind(cbind(fitted(max.IT.brm,re_formula = NA),forage.traits.max),
                      cbind(fitted(mean.IT.brm,re_formula = NA),forage.traits.mean))

#eusocial
eusocial.preds <- rbind(cbind(fitted(max.euc.phylo.reduced.brm,re_formula = NA),forage.traits.max),
                        cbind(fitted(mean.euc.phylo.reduced.brm,re_formula = NA),forage.traits.mean))

#physio vs functional
func.preds <- rbind(cbind(fitted(max.func.phylo.brm,re_formula = NA),forage.traits.max),
                        cbind(fitted(mean.func.phylo.brm,re_formula = NA),forage.traits.mean))

#overall trend
overall.plot <- ggplot(overall.pred,aes(y=(Estimate),x=ITD,fill="black")) +  
  xlab(NULL) + 
  ylab("ln Distance (m)") + 
  geom_ribbon(aes(ymin=(Q2.5),
                  ymax=(Q97.5)),
              alpha = 0.5,show.legend = FALSE,colour = NA)+
  geom_point(aes(y=log.dist),size=2.5,shape=21,
             show.legend = FALSE,alpha=0.6)+
  geom_line(size=1,show.legend = FALSE)+
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0,face="bold",size=14),
        legend.box.background = element_rect(colour = "black"),
        legend.title = element_text(size=14),
        legend.text = element_text(size=11),
        aspect.ratio = 1,
        strip.background = element_blank(),
        text=element_text(),
        axis.ticks.length = unit(0,"mm"),
        legend.title.align=0.5,
        axis.text.y = element_text(size=11),
        axis.text.x = element_text(size=11),
        axis.title.y = element_text(size=14),
        axis.title.x = element_text(size=14),
        strip.text = element_text(size=14,face = "bold"),
        panel.spacing = unit(0.5,"lines"),
        legend.background = element_blank(),
        panel.border = element_rect(color = "black",
                                    fill = NA, size = 0.4))+
  scale_fill_manual(values=col.palette) + 
  scale_colour_manual(values=col.palette) + 
  guides(linetype = FALSE)+
facet_wrap(~distance_type2)

func.preds$limits <- revalue(func.preds$limits,c("funct" = "Functional",
                                                 "physio" = "Physiological"))
#functional trend
func.plot <- ggplot(func.preds,aes(y=Estimate,x=ITD)) +  
  xlab("Body size (ITD, mm)") + 
  ylab("ln Distance (m)") + 
  geom_ribbon(aes(ymin=Q2.5,
                  ymax=Q97.5,col=limits,fill=limits),
              alpha = 0.5,show.legend = FALSE,colour = NA)+
  geom_point(aes(y=log.dist,fill=limits),col="black",size=2.5,shape=21,
             show.legend = TRUE,alpha=0.6)+
  geom_line(aes(col=limits,fill=limits),size=1,show.legend = TRUE)+
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0,face="bold",size=14),
        legend.box.background = element_rect(colour = "black"),
        legend.title = element_text(size=12,face="bold"),
        legend.text = element_text(size=11),
        aspect.ratio = 1,
        strip.background = element_blank(),
        text=element_text(),
        axis.ticks.length = unit(0,"mm"),
        legend.title.align=0.5,
        axis.text.y = element_text(size=11),
        axis.text.x = element_text(size=11),
        axis.title.y = element_text(size=14),
        axis.title.x = element_text(size=14),
        strip.text = element_blank(),
        panel.spacing = unit(0.5,"lines"),
        legend.background = element_blank(),
        panel.border = element_rect(color = "black",
                                    fill = NA, size = 0.4))+
  scale_fill_manual(values=col.palette[5:6]) + 
  scale_colour_manual(values=col.palette[5:6]) + 
  guides(linetype = FALSE)+
  labs(name="Foraging constraint",
       fill="Foraging constraint",
       col="Foraging constraint")+
  facet_wrap(~distance_type2)

two_plot <- overall.plot/func.plot
two_plot
```
**Figure 1:** Relationship between maximum (left) and typical (right) bee foraging distance (log-transformed metres) and foraging distances. Top: Overall trend, bottom: type of measurement used to estimate foraging distance: physiological (feeder or homing experiments), functional (other methods, see **Table 1**). Bold lines and ribbons indicate lines of best fit and corresponding 95% credible intervals. Circles denote raw data.

```{r Figure 2, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

#func eusocial
func.euc.preds <- rbind(cbind(fitted(max.func.euc.red.phylo.brm,re_formula = NA),forage.traits.max),
                    cbind(fitted(mean.func.euc.red.phylo.brm,re_formula = NA),forage.traits.mean))


col.palette <- c("#325756","#7d9fc2","#C582B2","#51806a","#4d5f8e","#A092B7")


#functional and eusocial trend
labels <- data.frame(limits=c("physio","physio","funct","funct"),
           distance_type2=c("Max","Typical","Max","Typical"),
           label=c("Max - Physiological","Typical - Physiological","Max - Functional","Max - Functional"),
           ITD=c(1,1,1,1),
           Estimate=c(11,11,11,11))

func.euc.preds$grp <- paste(func.euc.preds$distance_type2,func.euc.preds$social_tree)

func2.plot <- ggplot(func.euc.preds,aes(y=Estimate,x=ITD,col=social_tree,fill=social_tree)) +  
  xlab("Body size (ITD, mm)") + 
  ylab("ln Distance (m)") + 
  geom_ribbon(aes(ymin=Q2.5,
                  ymax=Q97.5),
              alpha = 0.5,show.legend = FALSE,colour = NA)+
  geom_point(aes(y=log.dist),col="black",size=2.5,shape=21,
             show.legend = TRUE,alpha=0.6)+
  geom_text(data=labels,aes(label = label,y=Estimate,x=ITD),inherit.aes = FALSE,hjust = 0,fontface="bold")+
  geom_line(aes(),size=1,show.legend = TRUE)+
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0,face="bold",size=14),
        legend.box.background = element_rect(colour = "black"),
        legend.title = element_text(size=12,face="bold"),
        legend.text = element_text(size=11),
        aspect.ratio = 1,
        strip.background = element_blank(),
        text=element_text(),
        axis.ticks.length = unit(0,"mm"),
        legend.title.align=0.5,
        axis.text.y = element_text(size=11),
        axis.text.x = element_text(size=11),
        axis.title.y = element_text(size=14),
        axis.title.x = element_text(size=14),
        strip.text = element_blank(),
        panel.spacing = unit(0.5,"lines"),
        legend.background = element_blank(),
        panel.border = element_rect(color = "black",
                                    fill = NA, size = 0.4))+
  scale_fill_manual(values=col.palette) + 
  scale_colour_manual(values=col.palette) + 
  guides(linetype = FALSE)+
  labs(name="Eusociality",
       fill="Eusociality",
       col="Eusociality")+
  facet_wrap(~limits+distance_type2)
func2.plot

```

**Figure 2:** Relationship between eusociality and phyisiological or functional bee foraging distances (log-transformed metres) and maximum (left) and typical (right) foraging distances. Top: Overall trend, bottom: type of measurement used to estimate foraging distance: physiological (feeder or homing experiments), functional (other methods, see **Table 1**). Bold lines and ribbons indicate lines of best fit and corresponding 95% credible intervals. Circles denote raw data.

##Q1 Overall trends - predicting maximum and typical foraging distances

```{r examples, echo=FALSE}
#standardised examples
#2mm 4mm bees
max.IT.fitted <- fitted(max.IT.brm,newdata=data.frame(ITD=c(2,4)),re_formula=NA)

max.IT.fitted[,1] <- exp(max.IT.fitted[,1])

mean.IT.fitted <- fitted(mean.IT.brm,newdata=data.frame(ITD=c(2,4)),re_formula=NA)

mean.IT.fitted[,1] <- exp(mean.IT.fitted[,1])


phys.preds <- data.frame(ITD=c(2,4,2,4),
           limits=c("physio","physio","funct","funct"))

max.IT.func.fitted <- cbind(phys.preds,fitted(max.func.phylo.brm,
                             newdata=phys.preds,
                             re_formula=NA))

mean.IT.func.fitted <- cbind(phys.preds,fitted(mean.func.phylo.brm,
                              newdata=phys.preds,
                              re_formula=NA))


max.IT.func.fitted[,3] <- exp(max.IT.func.fitted[,3])
mean.IT.func.fitted[,3] <- exp(mean.IT.func.fitted[,3])

#eusocial trends with physiological and functional
#3.30 (honeybee size)
euc.phys.preds <- data.frame(ITD=c(3.3,3.3,3.3,3.3,3.3,3.3),
                         limits=c("physio","physio","physio","funct","funct","funct"),
                         social_tree=c("Highly Eusocial","Primitively Eusocial","Solitary",
                                       "Highly Eusocial","Primitively Eusocial","Solitary"))
max.IT.euc.func.fitted <- cbind(euc.phys.preds,fitted(max.func.euc.red.phylo.brm,
                             newdata=euc.phys.preds,
                             re_formula=NA))
mean.IT.euc.func.fitted <- cbind(euc.phys.preds,fitted(mean.func.euc.red.phylo.brm,
                             newdata=euc.phys.preds,
                              re_formula=NA))

max.IT.euc.func.fitted[,4] <- exp(max.IT.euc.func.fitted[,4])
mean.IT.euc.func.fitted[,4] <- exp(mean.IT.euc.func.fitted[,4])
```
Overall, body size was positively related to both maximum (*Marginal R^2* = `r round(model.r2s[1,2][[1]],2)`) and typical (*Marginal R^2* = `r round(model.r2s[7,2][[1]],2)`) foraging distances (Figure 1). For example, the typical and maximum foraging distances of a middle sized bee (ITD: 4mm) are predicted to be `r max.IT.fitted[2,1]`m and `r mean.IT.fitted[2,1]`m, which are `r max.IT.fitted[2,1]-max.IT.fitted[1,1]`m and `r mean.IT.fitted[2,1]-mean.IT.fitted[1,1]`m longer than a small sized bee (ITD: 2mm, Max: `r max.IT.fitted[1,1]`m, Typ: `r mean.IT.fitted[1,1]`).

#Q2 Differences in constraints on foraging distances

We found that physiological and functional constraints/measurements upon foraging distances differ in their relationship (slope) with body size (Model table, Figure 2). Specifically, physiological measurements of foraging distances increased at a greater rate per unit ITD than functional constraints/measurements (significant interaction with ITD, steeper slope). For example,  in small bees (e.g. ITD of 2mm), physiological and functional measurements of foraging distance are similar (e.g. Typical: *functional*  `r mean.IT.func.fitted[3,3]`,*physiological* `r mean.IT.func.fitted[3,1]`;Maximum: *functional*  `r max.IT.func.fitted[3,3]`,*physiological* `r max.IT.func.fitted[3,1]`). However, in larger bees (ITD: 4mm), whereas physiological limits upon forging distance increase considerably, typical - `r mean.IT.func.fitted[3,2]`m, maximum - `r max.IT.func.fitted[3,2]`m, functional limits do not: typical, `r mean.IT.func.fitted[3,4]`m; maximum, `r max.IT.func.fitted[3,4]`m.

##Q2A) Effect of eusociality on physiological and functional bee foraging distances

Eusociality was highly important in structuring species differences in bee foraging distances (*Marginal R^2*: Typical distance: `r model.r2s[14,2]`, Maximum distance: `r model.r2s[6,2]`. In particular, highly eusocial bees (*Apis* species and stingless bees (Tribe: Meliponini)) exhibited significantly higher physiological and functional foraging distances than either primitively eusocial or solitary species (MODEL TABLE of "max.euc.phylo.reduced.brm"). Primitively eusocial and soitary species exhibited near-identical estimates of foraging distance. For example, a European honeybee *Apis mellifera* (ITD: `r print(traits[traits$Genus_species%in%"Apis_mellifera",]$ITD)`) is expected to have a physiological foraging distances of between  `r mean.IT.euc.func.fitted[1,4]` and `r max.IT.euc.func.fitted[1,4]`m. In contrast, a primitively eusocial or solitary bee of the same size, such as *Osmia pedicornis*, is expected to have a physiological foraging distances of between `r mean.IT.euc.func.fitted[3,4]` and `r max.IT.euc.func.fitted[3,4]`m.

**Table SX-XX. Model tables**
```{r, model tables, echo=FALSE}
summary(mean.IT.brm) #weak effect
summary(max.IT.brm)

summary(mean.func.phylo.brm)
summary(max.func.phylo.brm)

summary(mean.func.euc.red.phylo.brm) 
summary(max.func.euc.red.phylo.brm)

```



# Discussion
